Land cover mapping in data-sparse regions

### 2022-01-04 docs/nasa_meeting_poster.Rmd
Add meeting poster

---
title: "Manuscript figures"
author: "Lei Song"
date: "6/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Introduction

This is just a markdown to record the script to generate figures for the paper.

## Study area

Study area figure should include a overall view of study area in a bigger background, a zoom-in view of the study area (northern of Tanzania), the ecozones, the NICFI tiling system and the following sub-tiling system,

```{r}
library(here)
library(sf)
library(nngeo)
library(ggplot2)
library(ggspatial)
library(RColorBrewer)
library(dplyr)
library(rnaturalearth)

# Read boundaries
tz_bry <- read_sf(here('data/geoms/tanzania.geojson'))
study_area <- read_sf(here('data/geoms/study_area.geojson')) %>% 
    st_remove_holes()
nicfi_tiles <- read_sf(here('data/geoms/tiles_nicfi_north.geojson')) %>% 
    mutate(`NICFI tile` = 1)
ecozones <- read_sf(here('data/geoms/ecozones.geojson')) %>% 
    st_intersection(study_area) %>% 
    rename(Ecozone = zone)

# The overview of study area in the whole country at the left corner
# Get the surrounding countries
countries <- ne_countries(type = 'countries', scale = 'large',
                          returnclass = 'sf') %>% 
    st_intersection(., st_bbox(tz_bry) %>% 
                        st_as_sfc() %>% 
                        st_buffer(5, joinStyle = 'MITRE') %>% 
                        st_bbox(tz_bry) %>% # Get he squared one
                        st_as_sfc())

# Make overview
xlim <- c(st_bbox(countries)[1] + 3, st_bbox(countries)[3] - 3)
ylim <- c(st_bbox(countries)[2] + 3, st_bbox(countries)[4] - 3)
overview <- ggplotGrob(ggplot(tz_bry) +
    geom_sf(fill = 'lightblue', color = 'grey') +
    geom_sf_text(aes(label = geounit), size = 4) +
    geom_sf(data = countries, fill = NA, color = 'grey') +
    geom_sf(data = study_area, fill = 'red', color = 'grey') +
    coord_sf(xlim = xlim, ylim = ylim) +
    theme_bw() +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          plot.margin = unit(rep(0, 4), "cm")))

# Make the main figure
main_fg <- ggplot(ecozones) +
    geom_sf(aes(fill = Ecozone), 
            color = 'grey') +
    scale_fill_brewer(palette = 'Accent') +
    geom_sf(data = nicfi_tiles, fill = NA, 
            aes(color = "black", lty = "dotted"), 
            lwd = 0.4) +
    geom_sf(data = study_area, fill = NA, 
            aes(color = 'red', lty = 'solid')) +
    scale_color_identity(name = '',
                         breaks = c('black', 'red'),
                         labels = c('NICFI tiles', 'Study area'),
                         guide = 'legend') +
    scale_linetype_identity(name = '',
                            breaks = c('dotted', 'solid'),
                            labels = c('NICFI tiles', 'Study area'),
                            guide = 'legend') +
    annotation_custom(grob = overview, 
                      xmin = 36.5, xmax = 38.5,
                      ymin = -2.3, ymax = -0.5) +
    annotation_scale(location = 'br', text_cex = 1) +
    annotation_north_arrow(which_north = "true", location = 'tl') +
    theme_minimal() + 
    theme(axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          legend.text = element_text(size = 12),
          plot.margin = unit(rep(0, 4),"cm"))

ggsave(here('docs/figures/study_area.png'),
       main_fg,
       width = 10, height = 5.7, dpi = 600)
main_fg
```

## Satellite images

This figure will provide an example of satellite images including NICFI images with two seasons, and the harmonic regression of Sentinel-1 images.

```{r}
library(sf)
library(raster)
library(ggplot2)
library(ggpubr)
library(dplyr)
library(tidyr)
library(stringr)
library(RStoolbox)

# Read example images
# tile 1227-1002, 
# this tile has a few different land cover types and obvious characteristics
# so it could be a good example tile.
tile_id <- '1227-1002'
fnames_nicfi <- list.files(here('data/plts_nicfi'),
                           pattern = tile_id,
                           full.names = T)
nicfi_s1 <- stack(fnames_nicfi[1]) %>% subset(1:4)
nicfi_s2 <- stack(fnames_nicfi[2]) %>% subset(1:4)
rm(fnames_nicfi)

# Define example points
pts_example <- read_sf(here('data/north/pts_example.shp')) %>% 
    st_cast('POINT')

# NICFI
# Transform points
pts_nicfi <- pts_example %>% 
    st_transform(crs = crs(nicfi_s1))

# Define colors
cols_lc <- tibble(cols = c('#ff7f00', '#074702', '#b2df8a',
                           '#33a02c', '#3347f9', '#5a6057',
                           '#fdbf6f'),
                  `Land cover type` = c("Cropland", 'Forest', "Grassland", 
                                        'Shrubland',  "Water", "Built-up",
                                        'Bareland'),
                  ids = as.factor(1:7))

# Image of season 1
s1_fg <- ggRGB(nicfi_s1, r = 4, g = 3, b = 2, 
               scale = max(values(nicfi_s1)[, 2:4]), 
               stretch = 'hist') +
    geom_sf(data = pts_nicfi,
            color = 'black',
            size = 3) +
    geom_sf(data = pts_nicfi,
            aes(color = factor(id)),
            size = 2,
            show.legend = F) +
    scale_color_manual("Land cover type",
                       values = cols_lc$cols, 
                       breaks = cols_lc$ids,
                       labels = cols_lc$`Land cover type`) +
    theme_minimal() +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          plot.margin = unit(rep(0, 4),"cm"))

# Image of season 2
s2_fg <- ggRGB(nicfi_s2, r = 4, g = 3, b = 2, 
               scale = max(values(nicfi_s2)[, 2:4]), 
               stretch = 'lin') +
    geom_sf(data = pts_nicfi,
            color = 'black',
            size = 3) +
    geom_sf(data = pts_nicfi,
            aes(color = factor(id)),
            size = 2,
            show.legend = F) +
    scale_color_manual("Land cover type",
                       values = cols_lc$cols, 
                       breaks = cols_lc$ids,
                       labels = cols_lc$`Land cover type`) +
    theme_minimal() +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          plot.margin = unit(rep(0, 4),"cm"))

# Spectral signature of examples
# Extract values
pts_nicfi_s1 <- raster::extract(nicfi_s1, pts_nicfi) %>% 
    t() %>% data.frame() %>% 
    `names<-`(pts_nicfi$type) %>% 
    `rownames<-`(1:nrow(.)) %>% 
    mutate(`Band name` = c('Blue', 'Green', 'Red', 'NIR'),
           Band = 1:4) %>% 
    pivot_longer(cols = names(.)[1:(ncol(.) - 2)], 
                 names_to = 'Land cover type',
                 values_to = 'Reflectance') %>% 
    mutate(`Land cover type` = as.factor(`Land cover type`),
           Season = 'Season 1')
pts_nicfi_s2 <- raster::extract(nicfi_s2, pts_nicfi) %>% 
    t() %>% data.frame() %>% 
    `names<-`(pts_nicfi$type) %>% 
    `rownames<-`(1:nrow(.)) %>% 
    mutate(`Band name` = c('Blue', 'Green', 'Red', 'NIR'),
           Band = 1:4) %>% 
    pivot_longer(cols = names(.)[1:(ncol(.) - 2)], 
                 names_to = 'Land cover type',
                 values_to = 'Reflectance') %>% 
    mutate(`Land cover type` = as.factor(`Land cover type`),
           Season = 'Season 2')
pts_nicfi_all <- rbind(pts_nicfi_s1, pts_nicfi_s2)
rm(pts_nicfi_s1, pts_nicfi_s2)

# Figure
pts_nicfi_all <- merge(pts_nicfi_all, cols_lc[, 2:3])
ref_epl <- ggplot(pts_nicfi_all,
                  aes(x = Band, y = Reflectance / 1000,
                      color = ids)) +
    geom_point() + geom_line(show.legend = 'line') +
    ylab(bquote('Reflectance'~(10^3))) + 
    scale_color_manual("Land cover type",
                       values = cols_lc$cols, 
                       breaks = cols_lc$ids,
                       labels = cols_lc$`Land cover type`) +
    scale_x_discrete(limit = factor(1:4),
                     labels = unique(pts_nicfi_all$`Band name`)) +
    theme_bw() +
    theme(axis.title.x = element_blank()) +
    facet_wrap(~ Season, nrow = 2, 
               scales = "free_y",
               strip.position = "top")

nicfi_fg <- ggarrange(
    ggarrange(s1_fg, s2_fg,
              labels = c('Season 1', 'Season 2'),
              font.label = list(size = 12),
              vjust = 3,
              ncol = 2), 
    ref_epl, ncol = 2)
ggsave(here('docs/figures/nicfi_image.png'),
       nicfi_fg,
       width = 8, height = 2.7, dpi = 300)

# Sentinel-1 harmonic coefficient regression
fnames_s1 <- list.files(here('data/s1_harmonic'),
                        pattern = tile_id,
                        full.names = T)
vv_coefs <- stack(fnames_s1[str_detect(fnames_s1, 'VV')])
vh_coefs <- stack(fnames_s1[str_detect(fnames_s1, 'VH')])

# Image of VV
# Slope, Cos(x), Intercept
vv_fg <- ggRGB(vv_coefs, r = 2, g = 4, b = 1, 
               scale = max(values(vv_coefs)[, c(2, 4, 1)]), 
               stretch = 'hist') +
    geom_sf(data = pts_example,
            color = 'white',
            size = 3) +
    geom_sf(data = pts_example,
            aes(color = factor(id)),
            size = 2,
            show.legend = F) +
    scale_color_manual("Land cover type",
                       values = cols_lc$cols, 
                       breaks = cols_lc$ids,
                       labels = cols_lc$`Land cover type`) +
    theme_minimal() +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          plot.margin = unit(rep(0, 4),"cm"))

# Image of VH
# Slope, Cos(x), Intercept
vh_fg <- ggRGB(vh_coefs, r = 2, g = 4, b = 1, 
               scale = max(values(vh_coefs)[, c(1, 2, 4)]), 
               stretch = 'lin') +
    geom_sf(data = pts_example,
            color = 'white',
            size = 3) +
    geom_sf(data = pts_example,
            aes(color = factor(id)),
            size = 2,
            show.legend = F) +
    scale_color_manual("Land cover type",
                       values = cols_lc$cols, 
                       breaks = cols_lc$ids,
                       labels = cols_lc$`Land cover type`) +
    theme_minimal() +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          plot.margin = unit(rep(0, 4),"cm"))

# Time series and harmonic fitting of examples
# out = cos((2 * pi * n0 * t) / freq - (n % 2) * pi / 2)
# n_0 = {0,1,1,2,2}
# n = {0,1,2,3,4}
# freq = 365
# Band1: intercept
# Band2: slope
# Band3: sin(x)
# Band4: cos(x)
# Band5: sin(2x)
# Band6: cos(2x)

# Real time series
## Images
vv_ts <- stack(
    list.files(here('data/s1_timeseries_example/temp_ard_1227-1002'),
               pattern = 'VV',
               full.names = T)) 
vh_ts <- stack(
    list.files(here('data/s1_timeseries_example/temp_ard_1227-1002'),
               pattern = 'VH',
               full.names = T))

# DOY
doy <- c(0,   9,  12,  21,  33,  45,  48,  57,  69,  81,  93, 105, 117,
       129, 141, 153, 165, 177, 189, 201, 213, 225, 237, 249, 261, 273,
       285, 288, 293, 297, 300, 305, 309, 312, 317, 321, 324, 329, 333,
       336, 345, 348, 357, 360)

# Extract values
pts_vv_ts <- raster::extract(vv_ts, pts_example) %>% 
    t() %>% data.frame() %>% 
    `names<-`(pts_example$type) %>% 
    mutate(date = rownames(.)) %>% 
    `rownames<-`(1:nrow(.)) %>% 
    mutate(date = str_extract(date, '[0-9]{8}')) %>% 
    arrange(date) %>% 
    mutate(DOY = doy, polarization = 'VV',
           type = 'Real value')

pts_vh_ts <- raster::extract(vh_ts, pts_example) %>% 
    t() %>% data.frame() %>% 
    `names<-`(pts_example$type) %>% 
    mutate(date = rownames(.)) %>% 
    `rownames<-`(1:nrow(.)) %>% 
    mutate(date = str_extract(date, '[0-9]{8}')) %>% 
    arrange(date) %>% 
    mutate(DOY = doy, polarization = 'VH',
           type = 'Real value')

pts_ts <- rbind(pts_vv_ts, pts_vh_ts) %>% 
    dplyr::select(-date) %>% 
    pivot_longer(names(.)[1:7], 
                 names_to = "Land cover type", 
                 values_to = "dB")
rm(pts_vv_ts, pts_vh_ts)

# Harmonic coefficients
pts_vv <- raster::extract(vv_coefs, pts_example) %>% 
    t() %>% data.frame() %>% 
    `names<-`(pts_example$type) %>% 
    `rownames<-`(1:nrow(.)) %>% 
    mutate(`Band name` = c('Intersept', 'Slope', 
                           'Sin(x)', 'Cos(x)', 
                           'Sin(2x)', "Cos(2x)"), Band = 1:6)

pts_vv_fit <- do.call(cbind, lapply(names(pts_vv)[1:7], function(typ){
    coefs <- pts_vv %>% pull(typ)
    data.frame(val = coefs[1] + coefs[2] * doy +
    coefs[3] * sin(2 * pi * doy /365) + 
    coefs[4] * cos(2 * pi * doy /365) + 
    coefs[5] * sin(4 * pi * doy /365) + 
    coefs[6] * cos(4 * pi * doy /365))
})) %>% `names<-`(names(pts_vv)[1:7]) %>% 
    mutate(DOY = doy) %>% 
    mutate(polarization = 'VV',
           type = 'Harmoic regression')

pts_vh <- raster::extract(vh_coefs, pts_example) %>% 
    t() %>% data.frame() %>% 
    `names<-`(pts_example$type) %>% 
    `rownames<-`(1:nrow(.)) %>% 
    mutate(`Band name` = c('Intersept', 'Slope', 
                           'Sin(x)', 'Cos(x)', 
                           'Sin(2x)', "Cos(2x)"), Band = 1:6)

pts_vh_fit <- do.call(cbind, lapply(names(pts_vh)[1:7], function(typ){
    coefs <- pts_vh %>% pull(typ)
    data.frame(val = coefs[1] + coefs[2] * doy +
    coefs[3] * sin(2 * pi * doy /365) + 
    coefs[4] * cos(2 * pi * doy /365) + 
    coefs[5] * sin(4 * pi * doy /365) + 
    coefs[6] * cos(4 * pi * doy /365))
})) %>% `names<-`(names(pts_vh)[1:7]) %>% 
    mutate(DOY = doy) %>% 
    mutate(polarization = 'VH',
           type = 'Harmoic regression')

pts_fit <- rbind(pts_vv_fit, pts_vh_fit) %>% 
    pivot_longer(names(.)[1:7], 
                 names_to = "Land cover type", 
                 values_to = "dB")
rm(pts_vv_ts, pts_vh_ts)

# Make figure
pts_ts <- merge(pts_ts, cols_lc[, 2:3])
pts_fit <- merge(pts_fit, cols_lc[, 2:3])
ts_epl <- ggplot() + 
    geom_point(data = pts_ts, 
               aes(x = DOY, y = `dB`, 
                   color = factor(ids)),
               size = 0.6) +
    geom_line(data = pts_fit,
               aes(x = DOY, y = `dB`, 
                   color = factor(ids))) +
    facet_wrap(~polarization, 
               scales = 'free_y', nrow = 2,
               strip.position = "top") +
    scale_color_manual("Land cover type",
                       values = cols_lc$cols, 
                       breaks = cols_lc$ids,
                       labels = cols_lc$`Land cover type`) +
    theme_bw()

s1_fg <- ggarrange(
    ggarrange(vv_fg, vh_fg,
              labels = c('VV', 'VH'),
              font.label = list(size = 12),
              vjust = 3,
              ncol = 2), 
    ts_epl, ncol = 2)

ggsave(here('docs/figures/s1_image.png'),
       s1_fg,
       width = 8, height = 2.7, dpi = 300)
```

## Samples for Random forest and U-Net

This figure will provide the selected 200 tiles for RF to predict, the sub-tiling system for human checking, and the distribution of sub-tiles (chips) with score higher than 7.

```{r}
library(sf)
library(rasterVis)
library(ggnewscale)
library(cowplot)

@@

esb_fg <- ggarrange(plotlist = plots_esb, ncol = 4,
                    labels = sub_tiles$index, 
                    hjust = -2.5)
guess_fg <- ggarrange(plotlist = plots_guess, ncol = 4)
refine_fg <- ggarrange(plotlist = plots_refine, ncol = 4)
fgs <- ggarrange(esb_fg, guess_fg, refine_fg, nrow = 3,
                 labels = c('A', 'B', 'C'),
                 hjust = 0.4, vjust = 6,
                 font.label = list(face = "plain"))
labels_fg <- ggarrange(NULL, fgs, sub_tiling_fg, 
                       ncol = 3, widths = c(0.1, 2, 1))

ggsave(here('docs/figures/labels.png'),
       labels_fg,
       width = 8, height = 4, dpi = 300)
```

## Workflow

This figure will provide an overall flowchart of methods.

## Results
### Random forest evaluation
#### Holdout validate

```{r}
load(here("data/north/final_res.rda"))
cf <- confusionMatrix(data = final_res$.predictions[[1]]$.pred_class, 
                      reference = final_res$.predictions[[1]]$landcover)
save(cf, file = here('docs/figures/rf_holdout_conf.rda'))

conf <- cf$table
# row is truth, column is predict
# match with python sklearn
conf <- matrix(as.numeric(conf), nrow = nrow(conf), ncol = ncol(conf))
conf <- t(conf)

# Calculate producer's accuracy
prac <- sapply(1:7, function(n){
  conf[n, n] / rowSums(conf)[n]
})

acc_test <- data.frame(class = 1:7,
                       producer_acc = prac)
write.csv(acc_test, here('docs/figures/rf_holdout_acc.csv'),
          row.names = F)
```

#### Independent test

```{r}
library(terra)
library(parallel)
library(ramify)
library(ranger)
library(parsnip)
library(tidymodels)
library(caret)

tiles <- read_sf(here('data/geoms/tiles_nicfi_north.geojson'))
imgs <- rast(file.path(
        "/Volumes/elephant/pred_stack",
        paste0(tiles$tile[1], ".tif")))
sample_test <- read_sf(here('data/north/plys_holdout_check.geojson')) %>% 
  st_join(tiles) %>% st_transform(crs = crs(imgs)) %>% 
  vect(); rm(imgs)

test_rf_features <- do.call(
  rbind,
  mclapply(unique(sample_test$tile),
    function(tile_nm) {
      # Get imgs
      imgs <- rast(file.path(
        "/Volumes/elephant/pred_stack",
        paste0(tile_nm, ".tif")
      ))

      # get samples
      samples_this <- sample_test[sample_test$tile == tile_nm, ]
      trainings_this <- do.call(
          rbind, 
          lapply(1:nrow(samples_this), function(n){
          sample_n <- samples_this[n, ]
          terra::extract(
              imgs, sample_n) %>%
              mutate(landcover = sample_n$landcover) %>%
              dplyr::select(-ID)
      }))
      rm(imgs, samples_this)
      gc()
      trainings_this
    }, mc.cores = 4)
)
save(test_rf_features, 
     file = file.path(here("data/north"), 
                      "test_rf_features.rda"))

load(here("data/north/guess_rf_md.rda"))
test_rf_features <- test_rf_features %>% na.omit()
pred <- predict(guess_rf_md$fit, 
                test_rf_features %>% select(-landcover))
pred <- argmax(pred$predictions)
pred <- data.frame(predict = pred,
                   test = test_rf_features$landcover)
lc_types <- data.frame(id = 1:7,
                       type = c('cropland', 'forest', 'grassland',
                                'shrubland', 'water', 'built-up',
                                'bareland'))
pred_eval <- left_join(pred, lc_types, 
                       by = c('test' = 'type')) %>% 
    dplyr::select(-test) %>% 
    rename(test = id) %>% 
    mutate(predict = as.factor(predict),
           test = as.factor(test))

cf <- confusionMatrix(data = pred_eval$predict, 
                      reference = pred_eval$test)
save(cf, file = here('docs/figures/rf_test_conf.rda'))

acc_test <- cf$byClass
write.csv(acc_test, here('docs/figures/rf_test_acc.csv'),
          row.names = F)
```

#### Human checking
```{r}
library(sf)
library(glue)
library(here)

tile_nm <- 'catalog_sample_tiles_update.geojson'
tiles <- here(glue('results/north/{tile_nm}')) %>% 
  read_sf(); rm(tile_nm)
tiles <- tiles %>% 
  mutate(quality = glue("({score}, {hardiness})")) %>% 
  st_drop_geometry() %>% 
  mutate(quality = factor(quality))

ggplot(tiles, aes(quality)) +
  geom_bar(color = 'black',fill = 'cyan3') +
  xlab('Quality (correctness, difficulty)') +
  ylab('Count') +
  theme_bw()

quality_fig <- ggplot(
  tiles,
  aes(fill = factor(hardiness), x = score)) + 
  geom_bar(position = "stack") +
  scale_fill_brewer('Difficulty', palette = "Dark2") +
  xlab('Correctness') +
  ylab('Count') +
  theme_bw()

ggplot(
  tiles,
  aes(fill = factor(hardiness), x = score)) + 
  geom_bar(position = "stack") +
  coord_polar("x") +
  theme_bw()

ggsave(here('docs/figures/humancheck_quality.png'),
       quality_fig,
       width = 4, height = 4, dpi = 300)
```

### Overall results
#### Metrics
```{r}
library(glue)
library(stringr)
library(caret)

# Validate
# Get file names
fns <- list.files(here(
    'results/scalar/full_runs'), 
    pattern = 'unet_k15_only_nopad_clr_200epc',
    full.names = T)

# Metrics
fns <- fns[str_detect(fns, 'Validate')]
metrics_last10 <- do.call(cbind, lapply(fns, function(fn){
    read.csv(fn, stringsAsFactors = F) %>% 
        dplyr::select(-c(Wall.time, Step)) %>% 
        tail(10)
}))
names(metrics_last10) <- gsub(
    '.csv', "", 
    str_extract(fns, '[A-Z|a-z]{1}[A-Z|a-z]+.csv'))

# Calculate mean and sd
items <- c('loss', 'AA', 'mIoU', 'Cropland', 'Forest', 
           'Grassland', 'Shrubland', 'Water', 
           'Urban', 'Bareland')
types <- c('Cropland', 'Forest', 
           'Grassland', 'Shrubland', 'Water', 
           'Built-up', 'Bareland', 'AA')

tb <- do.call(rbind, lapply(items, function(item){
    vals <- metrics_last10 %>% pull(all_of(item))
    if (item == 'Urban') item <- 'Built-up'
    if (item == 'loss') item <- 'Loss'
    data.frame(mean = mean(vals),
               sd = sd(vals)) %>% 
        mutate(Metric = item)
}))

acc <- tb %>% filter(Metric %in% types) %>% 
    mutate(mean = round(mean * 100, 2),
           sd = round(sd * 100, 2),
           Class = factor(Metric, levels = types),
           Accuracy = glue(
               paste0('{format(round(mean, digits = 2), nsmall = 2)}',
                      '\u00B1{format(round(sd, digits = 2), nsmall = 2)} %'))) %>% 
    dplyr::select(Class, Accuracy) %>% 
    arrange(Class)
write.csv(acc, here('docs/figures/k15_only_acc_last10epoch.csv'),
          row.names = F)

# Test
sample_test <- read_sf(here('data/north/plys_holdout_check.geojson'))
pred <- rast(here(file.path('results/dl/prediction',
                            'unet_k15_only_clr_200epc',
                            'landcover_north.tif')))
sample_test <- vect(sample_test %>% st_transform(crs = crs(pred)))
pred_eval <- terra::extract(pred, sample_test, list = T)
pred_eval <- do.call(rbind, lapply(1:length(pred_eval), function(n){
    rst_vals <- pred_eval[[n]][[1]]
    data.frame(predict = rst_vals,
               test = sample_test[n, ]$landcover)
}))

lc_types <- data.frame(id = 1:7,
                       type = c('cropland', 'forest', 'grassland',
                                'shrubland', 'water', 'built-up',
                                'bareland'))
pred_eval <- left_join(pred_eval, lc_types, by = c('test' = 'type')) %>% 
    dplyr::select(-test) %>% 
    rename(test = id) %>% 
    mutate(predict = as.factor(predict),
           test = as.factor(test))
cf <- confusionMatrix(data=pred_eval$predict, reference = pred_eval$test)
save(cf, file = here('docs/figures/test_conf.rda'))

acc_test <- cf$byClass
write.csv(acc_test, here('docs/figures/test_acc.csv'),
          row.names = F)
```

#### Scalars

```{r}
library(here)
library(ggplot2)
library(ggpubr)

# Get file names
fns <- list.files(here(
    'results/scalar/full_runs'), 
    pattern = 'unet_k15_only_nopad_clr_200epc', 
    full.names = T)

step <- 59
# Learning rate
lr <- read.csv(fns[2], stringsAsFactors = F) %>% 
    mutate(Epoch = Step / step)

# Scalars
fns <- fns[str_detect(fns, 'Validate|Train_lr')]
final_mod_scalars <- lapply(fns, function(fn){
    scalar <- read.csv(fn, stringsAsFactors = F) %>% 
        mutate(Epoch = Step / step) %>% 
        dplyr::select(-c(Wall.time, Step))
})
names(final_mod_scalars) <- gsub(
    '.csv', "", 
    str_extract(fns, '[A-Z|a-z]{1}[A-Z|a-z]+.csv'))

types <- c('Learning rate', 'Loss', 'AA', 'mIoU', 'Cropland', 'Forest', 
           'Grassland', 'Shrubland', 'Water', 
           'Built-up', 'Bareland')
lc <- c('Cropland', 'Forest', 
        'Grassland', 'Shrubland', 'Water', 
        'Built-up', 'Bareland', 'AA')
content <- do.call(rbind, lapply(types, function(type){
    if (type == 'Built-up') {
        item <- 'Urban'
    } else if (type == 'Loss'){
        item <- 'loss'
    } else if (type == 'Learning rate'){
        item <- 'lr'
    } else {item <- type}
    final_mod_scalars[[item]] %>% 
        mutate(Item = type)
}))

>>>>

---
title: "Land cover mapping in data-sparse regions"
subtitle: Northern Tanzania
author: Lei Song, Lyndon Estes, Boka Luo, Anna Estes
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["default", "lucy", "middlebury-fonts", "theme.css"]
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(here)
```

```{r, eval = FALSE, include=FALSE}
library(here)
library(terra)
library(stars)
library(tmap)
library(ggpubr)
library(gridExtra)

pred_path <- here(file.path('results/dl/prediction',
                           'unet_k15_only_clr_200epc',
                           'landcover_north_withwl.tif'))
color_tb <- data.frame(
    id = 1:8,
    type = c("Cropland", "Forest", 
             "Grassland", "Shrubland", 
             "Water", 'Built-up', "Bareland", 
             'Wetland'),
    color = c('#ff7f00', '#074702', '#b2df8a',
              '#33a02c', '#3347f9', '#5a6057',
              '#fdbf6f', '#0ea591'))
# pred <- read_stars(pred_path)
pred <- rast(pred_path)
setCats(pred, value = color_tb$type, index = color_tb$id)
pred <- st_as_stars(pred)

# plot
tmap_mode("plot")
pred_final <- tm_shape(pred) +
    tm_raster("landcover_north", 
              title = "Land cover types",
              n = 8, style = 'cat',
              labels = color_tb$type,
              palette = color_tb$color) +
    tm_layout(frame = F,
              legend.show = F,
              outer.margins = c(0.02, 0.01, 0, 0.01),
              inner.margins = 0,
              bg.color = 'black')
pred_final <- tmap_grob(pred_final)

# Get legend
r_df <- data.frame(x = 1:8, 
                   y = 1:8,
                   rst = as.factor(1:8))
legend_lc <- ggplot() + 
    geom_raster(
        data = r_df, 
        aes(x = x, y = y, 
            fill = rst)) +
    scale_fill_manual('Land cover',
                      values = color_tb$color, 
                      breaks = 1:8,
                      labels = color_tb$type) +
    theme_minimal() +
    theme(legend.position = 'right',
          legend.text = element_text(color = 'white'),
          axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          plot.margin = unit(rep(0, 4),"cm"),
          plot.background = element_rect(fill = "black"))+
    guides(fill = guide_legend(ncol = 4))
legend_lc <- get_legend(legend_lc)
pred_final_plt <- ggarrange(pred_final, legend_lc, NA,
                        nrow = 3, heights = c(5, 1, 0.6))
ggsave(here('docs/figures/nasa_poster_result_img.png'),
       pred_final_plt, bg = 'black',
       width = 6, height = 6, dpi = 300)
ggsave(here('docs/figures/nasa_poster_title_img.png'),
       ggarrange(pred_final, bg = 'black',
       width = 6, height = 6, dpi = 300)
```

## Introductions

We proposed a method to make timely and accurate land cover maps in data-sparse regions, such as Tanzania. The main motivation of this project is that land cover products over Sub-Saharan Africa, including Tanzania, have poor quality.

For instance, the following is the evaluation of some LC products in Northern Tanzania based on our test dataset.

```{r}
library(tibble)
library(kableExtra)
lc_products <- tibble(
  `Product name` = c('CGLS_LC100m', 'FROM-GLC 2017v1', 'GFSAD30', 'TanSIS'),
  `Product type` = c('Land cover', 'Land cover', 'Crop mask', 'Crop mask'),
  Year = c(2018, 2017, 2015, 2018),
  Resolution = c('100 m', '30 m', '30 m', '250 m'),
  `Overall accuracy` = c('67.70%', '60.13%', '82.93%', '82.39%'))
lc_products %>%
  kbl() %>%
  kable_paper("striped", full_width = F) %>%
  row_spec(0, bold = T) %>% 
  row_spec(0:4, color = "white", background = 'black')
```

---

## Methodology

Stage 1: We extracted the consensus of multiple existing land cover products to generate fragmented pixel-wise training labels.

Stage 2: We translated pixel-wise training labels to image-wise labels using the random forest as a guessing model and satellite images.

Stage 3: We fine-tuned a U-Net network based upon these image-wise labels using Sentinel-1 time series and raw bands of NICFI basemaps as image features.

```{r, fig.align='center', out.width = "60%"}
knitr::include_graphics("figures/hrlcm_flowchart.png")
```

---

## Sub-tiling system and different types of label

```{r, eval=F}
library(raster)
library(glue)
library(here)
library(sf)
library(nngeo)
library(ggplot2)
library(RColorBrewer)
library(dplyr)
library(sf)
library(rasterVis)
library(ggnewscale)
library(cowplot)
library(ggpubr)

# Read boundaries
study_area <- read_sf(here('data/geoms/study_area.geojson')) %>% 
    st_remove_holes()
nicfi_tiles <- read_sf(here('data/geoms/tiles_nicfi_north.geojson')) %>% 
    mutate(`NICFI tile` = 1)

# Make the main figure
main_fg <- ggplot(nicfi_tiles) +
    geom_sf(fill = NA, 
            aes(color = "black", lty = "dotted"), 
            lwd = 0.4) +
    geom_sf(data = study_area, fill = NA, 
            aes(color = 'red', lty = 'solid')) +
  scale_color_identity(name = '',
                         breaks = c('black', 'red'),
                         labels = c('NICFI tiles', 'Study area'),
                         guide = 'legend') +
    scale_linetype_identity(name = '',
                            breaks = c('dotted', 'solid'),
                            labels = c('NICFI tiles', 'Study area'),
                            guide = 'legend') +
    theme_minimal() + 
    theme(legend.position = 'none',
          axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          plot.margin = unit(rep(0, 4),"cm"))
save(main_fg, file = 'figures/nasa_tiling.rda')

@@

esb_fg <- ggarrange(plotlist = plots_esb, ncol = 4,
                    labels = sub_tiles$index, 
                    hjust = -2.5)
guess_fg <- ggarrange(plotlist = plots_guess, ncol = 4)
refine_fg <- ggarrange(plotlist = plots_refine, ncol = 4)
fgs <- ggarrange(esb_fg, guess_fg, refine_fg, nrow = 3,
                 labels = c('A', 'B', 'C'),
                 hjust = 0.4, vjust = 6,
                 font.label = list(face = "plain"))
labels_fg <- ggarrange(NULL, fgs, sub_tiling_fg, 
                       ncol = 3, widths = c(0.1, 2, 1))
fg <- ggarrange(main_fg, labels_fg, ncol = 2, widths = c(1, 1.8))
ggsave(here('docs/figures/nasa_poster_exp_img.png'),
       fg, bg = 'white', width = 10, height = 4, dpi = 300)
```

- We used NICFI basemap tiling system to manage the whole study area. 
- We split each tile into 8 * 8 sub-tiles. Within each tile, we randomly selected 4 sub-tiles (3 for training and 1 for validation) to make image-wise training labels with quality information.

```{r, fig.align='center', warning=F, message=F, fig.width=10, fig.height=4, fig.cap = "A: fragmented consensus labels, B: Image-wise RF guessed label, C: Human refined label"}
knitr::include_graphics("figures/nasa_poster_exp_img.png")
```

---

## U-Net training

```{r, fig.align='center', out.width = "90%"}
knitr::include_graphics("figures/curves_k15_only_final_mod.png")
```
---

## Results

```{r, fig.align='center', out.width = "80%"}
knitr::include_graphics("figures/nasa_poster_result_img.png")
```

---

## Results

```{r, fig.align='center', out.width = "90%", warning=F, message=F}
knitr::include_graphics("figures/nasa_poster_eval.png")
load(here('docs/figures/conf_mean_std.rda'))
mean_std %>% 
  dplyr::filter(Landcover != 'Wetland') %>% 
  t() %>% 
  kbl() %>%
  kable_paper("striped", full_width = F, font_size = 16) %>%
  column_spec(1, bold = T) %>% 
  row_spec(1:2, color = "white", background = 'black')
```
